\section{Testing Determinism with OMNeT++
Fingerprints}\label{sec:deterministic-tests}

Deterministic tests involve multiple runs with identical configurations and
random number generator seeds, verifying that each simulation run yields the
\textit{exact} same results, a critical requirement for reproducibility in
simulations.

For this test, a configuration of 10 nodes was used, including three miners,
with each node generating an average of one transaction every 10 seconds.

This test leverages an \omnetpp{} feature that computes a simulation
``fingerprint'', a hash updated throughout the simulation based on various
event properties, and displayed at the simulation's conclusion. \omnetpp{} then
compares the final fingerprint with a predefined expected value specified in
the configuration file. Ingredients used in the fingerprint calculation can be
set in the configuration file \cite[Section~15.4]{omnetpp-simulation-manual}.

All statistics collection was enabled, and each run simulated 24 hours.

In the initial run to generate the fingerprint, the following components were
included for calculation:
\begin{multicols}{2}
	\begin{itemize}
		\item The event number;
		\item The simulation time;
		\item The message/event full name;
		\item The message (packet) bit length;
		\item The module full path;
		\item The random numbers drawn;
		\item The scalar results;
		\item The statistic results (histograms, etc.);
		\item The vector results.
	\end{itemize}
\end{multicols}

These components were selected to capture key properties of each simulation
aspect and object. The last three items --- scalar results, statistical results,
and vector results --- are especially critical for ensuring reproducibility.
Anyway, also other ingredients were included in order to guarantee the
simulation's trajectory is always \textit{exactly} the same.

The generated fingerprint was:
\begin{verbatim}
5250-e519/etnlprszv
\end{verbatim}

With this fingerprint set in the configuration, the simulation was repeated 100
times. All 100 runs successfully matched the expected fingerprint, with each
output file containing the ``Fingerprint successfully verified'' message:
\begin{verbatim}
$ grep -c 'Fingerprint successfully verified' *.out | wc -l
100
\end{verbatim}
