\chapter{Future Work}\label{ch:next}

While \iblock{} successfully simulates the construction and propagation of
blocks and transactions in a Bitcoin network, the Bitcoin protocol encompasses
far more than these basic operations. Expanding \iblock{} to incorporate
additional protocol features would enhance its utility, enabling a broader
range of experiments and deeper insights into Bitcoin and blockchain dynamics.

The most critical feature currently absent in \iblock{} is the implementation
of Bitcoin's \textbf{gossip protocol}. In a real Bitcoin network, each node
connects to a limited number of peers. When a node receives a block or
transaction, it notifies its peers via \emph{inventory messages}. These peers
can then decide whether to request the data using a \code{GETDATA} message,
depending on whether they already possess it or not. This propagation process
continues until the entire network is updated. Incorporating this mechanism
would provide a more accurate simulation of Bitcoin's peer-to-peer
communication and its potential bottlenecks or inefficiencies. Furthermore,
implementing the gossip protocol would enable the study of other types of
attacks, such as the \emph{eclipse attack}, where an attacker isolates a node
by controlling its connections and feeds it with false informations to perform,
for example, a double-spending attack against the victim \cite{eclipse}.

Beyond the gossip protocol, other key Bitcoin mechanisms remain to be
implemented, including:
\begin{description}
	\item[Peer discovery] Nodes query peers for lists of other nodes to
		expand their connections;
	\item[Bloom filters] Utilized by Simplified Payment Verification (SPV)
		nodes to query the blockchain for specific transactions while
		maintaining user privacy;
	\item[Segregated Witness (SegWit)] A protocol improvement that
		separates transaction signatures from the transaction data to
		optimize storage and reduce block sizes (and thus fees);
	\item[Other Protocol Features] Elements like transaction malleability
		prevention, multi-signature schemes, and advanced scripting
		capabilities are essential for comprehensive protocol support.
\end{description}
A complete Bitcoin simulator would integrate these features, providing the
capability to study nuanced protocol behaviors and optimizations.

The next logical step in \iblock{}'s evolution is to implement support for
\omnetpp{}'s links. This addition would enable the simulation of a more
realistic network environment, where nodes communicate over channels with
varying bandwidths and latencies. Such functionality is critical for modeling
real-world network conditions and is a prerequisite for accurately simulating
the gossip protocol. Although partial progress has been made in this area, as
discussed in \secref{sec:partial-work}, completing this feature remains a
priority.

Looking forward, \iblock{} could be extended to simulate other blockchain
systems, such as Ethereum \cite{ethereum}. Ethereum introduces unique concepts
like Proof-of-Stake, smart contracts, gas mechanisms, and an account-based
model, which differ significantly from Bitcoin's UTXO-based approach.
Developing a more generic framework capable of supporting multiple blockchain
systems would be a significant milestone. Such a framework could allow
researchers and developers to simulate diverse blockchain ecosystems under a
unified architecture, broadening \iblock{}'s applicability.

By incorporating these enhancements, \iblock{} can evolve into a robust,
versatile simulator for blockchain research. Features such as the gossip
protocol, peer discovery, and SegWit, combined with support for \omnetpp{}
links and channels, will deepen the simulation's fidelity. Furthermore,
extending the framework to support systems like Ethereum would position
\iblock{} as a universal tool for studying blockchain-based networks, unlocking
new opportunities for research and development.

\input{next/inet}
