\chapter{Testing the Simulation Model}\label{ch:testing}

This chapter outlines the tests conducted to verify the simulation model.

Two types of tests should be perfomed on every newly developed simulation
model: verification and validation tests. Verification and validation are
distinct but complementary processes critical to building confidence in the
model's reliability and effectiveness. Verification confirms that the model
accurately implements the conceptual design, while validation assesses the
model's ability to represent real-world system behavior
\cite[Chapter~15]{omnetpp-simulation-manual}.

Verification is the process of ensuring that the simulation model has been
correctly built in accordance with the conceptual model. This phase is focused
on detecting and eliminating any discrepancies between the implemented model
and its initial design, concentrating on accuracy in logic, consistency in data
flows, and correctness of algorithms. In this project, verification tests were
designed to catch issues such as programming errors, incorrect parameter
settings, and unintended exceptions that might disrupt the simulation.

Verification thus serves as an internal consistency check, confirming that the
logic and processes coded in the model adhere strictly to the specifications of
the conceptual model.

On the other hand, validation is the process of determining whether the
simulation model accurately represents the actual system it is meant to
emulate. Unlike verification, which focuses on the internal consistency of the
model with its design, validation is an external assessment that compares the
model's outputs to real-world data or expected behavior patterns. In this
phase, tests are conducted to evaluate whether the model produces results that
are consistent with observed data or accepted theoretical outcomes of the real
system.

Validation tests are essential for ensuring that the simulation's predictions,
trends, and responses accurately reflect what would be observed in the real
environment. Methods used in validation can include comparing simulated outputs
to empirical data, using statistical tests to check for significant deviations,
and conducting sensitivity analyses to see if the model behaves consistently
when variables are adjusted. Validation provides confidence that the simulation
model is a reliable tool for making predictions and deriving insights about the
real system.

Validation of the model will be done using the experiments described in
\chref{ch:pocs}, by checking that the results of the simulations are consistent
with theoretical and empirical data drawn from the real Bitcoin network and
from other related works. This chapter focuses on the verification tests
conducted on \iblock{}.

The configurations for the verification tests presented in this
chapter are available in the \texttt{simulations/tests.ini} file of the
repository.

\input{testing/manual}
\input{testing/smoke}
\input{testing/leaks}
\input{testing/fingerprints}
